---
title: Data Types
author: R package build
date: '2022-02-21'
slug: data-types
categories: []
tags: []
---

```{r}
library(tidyverse)
```
# Cat lovers
```{r}
urlcat <- 'https://raw.githubusercontent.com/gnunez2020/tidyverse/main/cat-lovers.csv'
cat_lovers <- read_csv(urlcat)
names(cat_lovers)
```




# Warnings
```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats))
```
```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats, na.rm = TRUE))
```

# Fix
```{r}
glimpse(cat_lovers)
```


```{r}
cat_lovers %>%
  mutate(number_of_cats = case_when(
    name == "Ginger Clark" ~ 2,
    name == "Doug Bass"    ~ 3,
    TRUE                   ~ as.numeric(number_of_cats)
    )) %>%
  summarise(mean_cats = mean(number_of_cats))
```

```{r}
cat_lovers %>%
  mutate(
    number_of_cats = case_when(
      name == "Ginger Clark" ~ "2",
      name == "Doug Bass"    ~ "3",
      TRUE                   ~ number_of_cats
      ),
    number_of_cats = as.numeric(number_of_cats)
    ) %>%
  summarise(mean_cats = mean(number_of_cats))
```

# Data Types in R

## logical
### boolean values TRUE and FALSE
```{r}
typeof(TRUE)
```

## double
### floating point numerical values (default numerical type)
```{r}
typeof(1.335)

typeof(7)
```

## integer
### integer numerical values (indicated with an L)
```{r}
typeof(7L)

typeof(1:3)
```

## character
### character strings
```{r}
typeof("hello")
```

# Concatenation
## Vectors can be constructed using the c() function.
```{r}
c(1, 2, 3)

c("Hello", "World!")

c(c("hi", "hello"), c("bye", "jello"))
```

# Converting between types
```{r}
x <- 1:3
x

typeof(x)

```

```{r}
y <- as.character(x)
y

typeof(y)

```

```{r}

x <- c(TRUE, FALSE)
x

typeof(x)
```

```{r}
y <- as.numeric(x)
y

typeof(y)
```

## R will happily convert between various types without complaint when different types of data are concatenated in a vector, and that's not always a great thing!

```{r}
c(1, "Hello")

c(FALSE, 3L)
```

```{r}
c(1.2, 3L)

c(2L, "two")
```

# Explicit vs. implicit coercion
## Explicit coercion is when you call a function like as.logical(), as.numeric(), as.integer(), as.double(), or as.character()

## Implicit coercion happens when you use a vector in a specific context that expects a certain type of vector



# Special values
## NA: Not available
## NaN: Not a number
## Inf: Positive infinity
## -Inf: Negative infinity

```{r}
pi / 0
0 / 0
1/0 - 1/0
1/0 + 1/0
```

## Special
```{r}
x <- c(1, 2, 3, 4, NA)
mean(x)
mean(x, na.rm = TRUE)
summary(x)
```

## Logical
```{r}
typeof(NA)
```

# Mental model for NAs
## Unlike NaN, NAs are genuinely unknown values
## But that doesn't mean they can't function in a logical way
```{r}
# TRUE or NA
TRUE | NA

# FALSE or NA
FALSE | NA
```
## NA is unknown, so it could be TRUE or FALSE

### TRUE | NA
```{r}
TRUE | TRUE  # if NA was TRUE

TRUE | FALSE # if NA was FALSE
```

### FALSE | NA
```{r}
FALSE | TRUE  # if NA was TRUE

FALSE | FALSE # if NA was FALSE

```

