---
title: Importing data
author: R package build
date: '2022-02-22'
slug: importing-data
categories: []
tags: []
---
```{r}
library(tidyverse)
```

# readr
>
read_csv() - comma delimited files
read_csv2() - semicolon separated files (common in countries where , is used as the decimal place)
read_tsv() - tab delimited files
read_delim() - reads in files with any delimiter
read_fwf() - fixed width files

# readxl
>
read_excel() - read xls or xlsx files

# Reading data
```{r}
nobel <- read_csv(file = "https://raw.githubusercontent.com/gnunez2020/tidyverse/main/nobel.csv")
nobel
```

# Writing data
```{r}
# Write a file
df <- tribble(
  ~x, ~y,
  1,  "a",
  2,  "b",
  3,  "c"
)
df
```

# Variable names

## Data with bad names
```{r}
edibnb_badnames <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/edibnb-badnames.csv")
names(edibnb_badnames)
```
#... but R doesn't allow spaces in variable names


### Option 1 - Define column names
```{r}
edibnb_col_names <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/edibnb-badnames.csv",
                             col_names = c("id", "price", 
                                           "neighbourhood", "accommodates",
                                           "bathroom", "bedroom", 
                                           "bed", "review_scores_rating", 
                                           "n_reviews", "url"))
names(edibnb_col_names)
```

### Option 2 - Format text to snake_case
```{r}
edibnb_clean_names <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/edibnb-badnames.csv") %>%
  janitor::clean_names()
names(edibnb_clean_names)
```

# Variable types
```{r}
read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/df-na.csv")
# Which type is x? Why?
```
## Option 1. Explicit NAs
```{r}
read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/df-na.csv", 
         na = c("", "NA", ".", "9999", "Not applicable"))
```

## Option 2. Specify column types
```{r}
read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/df-na.csv", col_types = list(col_double(), 
                                            col_character(), 
                                            col_character()))
```

# Column Types
```{r}
read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/df-na.csv")
```

# Case study: Favourite foods
```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/favourite-food.csv")
fav_food
```

## Variable names
```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/favourite-food.csv") %>%
  janitor::clean_names()
fav_food
```

## Handling NAs
```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/favourite-food.csv",
  na = c("N/A", "99999")) %>%
  janitor::clean_names()
fav_food
```

## Make age numeric
```{r}
fav_food <- fav_food %>%
  mutate(
    age = if_else(age == "five", "5", age),
    age = as.numeric(age)
    )
glimpse(fav_food)
```

## Socio-economic status
```{r}
fav_food %>%
  count(ses)
```

## Make ses factor
```{r}
fav_food <- fav_food %>%
  mutate(ses = fct_relevel(ses, "Low", "Middle", "High"))
fav_food %>%
  count(ses)
```

## Putting it altogether
```{r}
fav_food <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/favourite-food.csv", na = c("N/A", "99999")) %>%
  janitor::clean_names() %>%
  mutate(
    age = if_else(age == "five", "5", age), 
    age = as.numeric(age),
    ses = fct_relevel(ses, "Low", "Middle", "High")
  )
fav_food
```

## Out and back in
```{r}
fav_food_clean <- read_csv("https://raw.githubusercontent.com/gnunez2020/tidyverse/main/favourite-food.csv")
```

# read_rds() and write_rds()
>
CSVs can be unreliable for saving interim results if there is specific variable type information you want to hold on to.
An alternative is RDS files, you can read and write them with read_rds() and write_rds(), respectively.
>
read_rds(path)
write_rds(x, path)


# Other types of data
> 
googlesheets4: Google Sheets
haven: SPSS, Stata, and SAS files
DBI, along with a database specific backend (e.g. RMySQL, RSQLite, RPostgreSQL etc): allows you to run SQL queries against a database and return a data frame
jsonline: JSON
xml2: xml
rvest: web scraping
httr: web APIs
sparklyr: data loaded into spark


