---
title: Single data frame Operations
author: R package build
date: '2022-02-14'
slug: single-data-frame-operations
categories: []
tags: []
---

## Call libraries
```{r}
library(tidyverse)
```

##Select

### Select to include

```{r}
hotels %>%
  select(hotel, lead_time)
```

### Select to exclude

```{r}
hotels %>%
  select(-agent)
```

### Select a range
```{r}
hotels %>%
  select(hotel:arrival_date_month)

hotels %>%
  select(ends_with("type"))
```

### Select a characteristic

```{r}
hotels %>%
  select(starts_with("arrival"))
```

### Select ascending/decending

```{r}

hotels %>%
  select(adults, children, babies) %>%
  arrange(babies)

hotels %>%
  select(adults, children, babies) %>%
  arrange(desc(babies))
```

## Slice 

###Slice for first five rows

```{r}
hotels %>%
  slice(1:5)
```

## Filter

### Filter for subset of rows
```{r}
hotels %>%
  filter(hotel == "City Hotel")
```

### Filter for many conditions

```{r}
hotels %>%
  filter( 
    adults == 0,
    children >= 1
    ) %>% 
  select(adults, babies, children)
```
### Filter for complex conditions

```{r}
hotels %>%
  filter( 
    adults == 0,     
    children >= 1 | babies >= 1     # | means or
    ) %>%
  select(adults, babies, children)
```

## Distinct and Count

### Distinct to filter for unique rows
```{r}
hotels %>% 
  distinct(hotel, market_segment) %>%
  arrange(hotel, market_segment)
```

### Count to create frequency tables and arrange

```{r}

# alphabetical order by default
hotels %>%
  count(market_segment) %>%
  arrange(n)

# descending frequency order
hotels %>%
  count(market_segment, sort = TRUE) %>%
  arrange(desc(n))
```

### Count for multiple variables

```{r}
hotels %>%
  count(hotel, market_segment)
```

### order matters when you count

```{r}
# hotel type first
hotels %>%
  count(hotel, market_segment)

# market segment first
hotels %>%
  count(market_segment, hotel)
```

## Mutate

### Mutate to add a new variable

```{r}
hotels %>%
  mutate(little_ones = children + babies) %>%
  select(children, babies, little_ones) %>%
  arrange(desc(little_ones))

# Resort Hotel
hotels %>%
  mutate(little_ones = children + babies) %>%
  filter(
    little_ones >= 1,
    hotel == "Resort Hotel"
    ) %>%
  select(hotel, little_ones)

# City Hotel
hotels %>%
  mutate(little_ones = children + babies) %>%
  filter(
    little_ones >= 1,
    hotel == "City Hotel"
    )  %>%
  select(hotel, little_ones)
```

## Summarise and Group_by

### Summarise for summary stats

```{r}
# mean average daily rate for all bookings
hotels %>%
  summarise(mean_adr = mean(adr))
```
### Group_by for grouped operations

```{r}
# mean average daily rate for all booking at city and resort hotels
hotels %>%
  group_by(hotel) %>%
  summarise(mean_adr = mean(adr))
```

### Calculating frequencies
```{r}
# The following two give the same result, so count is simply short for group_by then determine frequencies

hotels %>%
  group_by(hotel) %>%
  summarise(n = n())

hotels %>%
  count(hotel)
```

### Multiple summary statistics

```{r}
hotels %>%
  summarise(
    min_adr = min(adr),
    mean_adr = mean(adr),
    median_adr = median(adr),
    max_adr = max(adr)
    )
```

