---
title: Data Classes
author: R package build
date: '2022-02-21'
slug: data-classes
categories: []
tags: []
---

```{r}
library(tidyverse)
```
# Cat lovers
```{r}
urlcat <- 'https://raw.githubusercontent.com/gnunez2020/tidyverse/main/cat-lovers.csv'
cat_lovers <- read_csv(urlcat)
names(cat_lovers)
```
# Hotels
```{r}
urlhotel <- 'https://raw.githubusercontent.com/gnunez2020/tidyverse/main/hotels.csv'
hotels <- read_csv(urlhotel)
names(hotels)
```

# Classes
##The class attribute relates to the S3 class of an object which determines its behaviour
###Examples: factors, dates, and data frames

# Factors
##R uses factors to handle categorical variables, variables that have a fixed and known set of possible values
```{r}
x <- factor(c("BS", "MS", "PhD", "MS"))
x

typeof(x)
class(x)
glimpse(x)
as.integer(x)

```

# Dates
##We can think of dates like an integer (the number of days since the origin, 1 Jan 1970) and an integer (the origin) glued together
```{r}
y <- as.Date("2020-01-01")
y

typeof(y)
class(y)
as.integer(y)
as.integer(y) / 365 # roughly 50 yrs

```

# Data frames
##We can think of data frames like like vectors of equal length glued together
```{r}
df <- data.frame(x = 1:2, y = 3:4)
df
typeof(df)
class(df)

```

# Lists
##Lists are a generic vector container vectors of any type can go in them
```{r}
l <- list(
  x = 1:4,
  y = c("hi", "hello", "jello"),
  z = c(TRUE, FALSE)
)
l
```

# Lists and data frames
##A data frame is a special list containing vectors of equal length
##When we use the pull() function, we extract a vector from the data frame
```{r}
df
df %>%
  pull(y)
```
# Working with factors

# Read data in as character strings
```{r}
glimpse(cat_lovers)

ggplot(cat_lovers, mapping = aes(x = handedness)) +
  geom_bar()

cat_lovers %>%
  mutate(handedness = fct_infreq(handedness)) %>%
  ggplot(mapping = aes(x = handedness)) +
  geom_bar()
```

# Come for the functionality
##Factors are useful when you have true categorical data and you want to override the ordering of character vectors to improve display
##They are also useful in modeling scenarios
##The forcats package provides a suite of useful tools that solve common problems with factors

# Working with dates
##lubridate is the tidyverse-friendly package that makes dealing with dates a little easier
##It's not one of the core tidyverse packages, hence it's installed with install.packages("tidyverse) but it's not loaded with it, and needs to be explicitly loaded with library(lubridate)

```{r}
hotels %>%
  select(starts_with("arrival_"))
# Calculate and visualise the number of bookings on any given arrival date.
```

## Step 1. Construct dates
```{r}
library(glue)
hotels %>%
  mutate(
    arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")
    ) %>% 
  relocate(arrival_date)
```

## Step 2. Count bookings per date
```{r}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date)
```

## Step 3. Visualise bookings per date
```{r}
hotels %>%
  mutate(arrival_date = glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}")) %>%
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

## Error
##Why does the plot start with August when we know our data start in July? And why does 10 August come after 1 August?

## Step 1. REVISED Construct dates "as dates"
```{r}
library(lubridate)
hotels %>%
  mutate(
    arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))
    ) %>% 
  relocate(arrival_date)
```

## Step 2. Count bookings per date
```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date)
```

## Step 3a. Visualise bookings per date
```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_line()
```

## Step 3b. Visualise using a smooth curve
```{r}
hotels %>%
  mutate(arrival_date = ymd(glue("{arrival_date_year} {arrival_date_month} {arrival_date_day_of_month}"))) %>% 
  count(arrival_date) %>%
  ggplot(aes(x = arrival_date, y = n, group = 1)) +
  geom_smooth()
```

